### 2022-05-29 article.tex
Progress

@@

@@

To **obtain** the genetic data, yes, there may be fieldwork and/or lab-work
involved. For example, a researcher needs to go into the field to take
samples (e.g. blood) of humans or other species. To analyse these samples,
a researcher needs a lab. Depending on the technique being used on
how to analyse the samples, there may be a big bioinformatics step
to aggregate the measurements into useful genetic data. 
Although this paper does not focus on the code run to 
aggragate the raw data into useful genetic data, the same arguments
can be made for that code as in the example: that code is 
the ground truth of how the raw data is collected into useful genetic data.

@@

\paragraph{Code itself has important properties}

The code of a computational experiment is
more important to preserve than the article describing
the experiment, even when both describe the experiment identically,
correctly and reproducibly. 
Here I will argue why this is the
case and the implications and best practices for knowledge preservation.
To summarize, (1) it gives an honest indication of the quality 
of a computational experiment, ....

Optionally, code can actively teach what it does.
Here, I will go into more detail of each point.

\paragraph{Example of code being convincing}

As code is paradata, as it describes how data is being collected,
it can convince (or not) that it does so correctly.
Similar to a cell biologist that can be working sterile to avoid contamination
by airborne bacteria, 
or sloppy and likelier to have contaminated samples,
code can be examplary or sloppy.

@@

 * Number of commits
 * Commit history
 * Code coverage
 * Objective indicator of work being done

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Example study}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Code is more important than the (English) paper}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

To \emph{obtain} the genetic data, yes, there may be fieldwork and/or lab-work
involved. For example, a researcher needs to go into the field to take
samples (e.g. blood) of humans or other species. To analyse these samples,
a researcher needs a lab. Depending on the technique being used on
how to analyse the samples, there may be a big bioinformatics step
to aggregate the measurements into useful genetic data. 
Although this paper does not focus on the code run to 
aggragate the raw data into useful genetic data, the same arguments
can be made for that code as in the example: that code is 
the ground truth of how the raw data is collected into useful genetic data.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{There are multiple way to publish code}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

There are multiple ways of publishing code,
that are increasingly hard to do, yet result
in more reproducible experiments:
(1) publishing the code as text only, e.g. by printing
it in a Supplementary Materials, (2) publishing
the code on a hosted website, e.g on GitHub, with
added features (3) publishing 
the installed code within a virtual environment,
for example, using a Singularity container.
Here I will go into details of each of these ways.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Publishing raw code}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Code is commonly published alongside of a 
paper, pasted as text in the Supplementary Materials 
(for example, in \cite{labrecque2019interpretation}).
This is the rawest form of publishing code,
as it is just text.

\paragraph{Raw code is important}

The raw code of a computational experiment is
more important to preserve than the article describing
the experiment, even when both describe the experiment identically,
correctly and reproducibly, 
as it has/should been that code that generated the results of an experiment.
Again: if the paper and code differ, it is the code that did
the actual work with the actual/true results.

\paragraph{Raw code is an indicator of quality}

Raw code gives an indication of the quality 
of a computational experiment and several metrics are
devised to get an idea of the quality of code.

The simplest metric to assess the quality of code is 
the amount of (single) lines of code (SLOC). 
Code that has a lower SLOC count, i.e. it is more concise,
is usually regarded as having a lower chance of containing bugs.

A more interesting metric is the cyclomatic complexity of the code.
The cyclomatic complexity is defined as the number of ways that
the code can be executed. 
For example, code that has only one \verb|if| statement
has a cyclomatic complexity of 2, as the condition within the \verb|if|
statement can be true or false.
The cyclomatic complexity correlates with code complexity,
where more complex code is likelier to contain bugs.

A feature that deserves more attention
is that code can actively teach what it does.
A mechanism to do so, which is present in most programming
languages, is error handling: errors are the mechanism
by which a programmer can teach the user what to do/expect.
One example is when calculating the variance of a distribution.
To calculate the variance of a distribution, at least two values
are required. When the variance of only one value is requested,
an error message can helpfully indicate that one needs at least two values.

\paragraph{The problems with raw code}

There are multiple problems with publishing raw code.
First, there is no guarantee that that code \emph{can} actually
run, for example, due to a typo added after the code was pasted.
Secondly, there is no way to determine if code that worked yesterday
works today. These problems can be prevented by, instead of
pasting raw code, using a website to host that code (see below).

\paragraph{Raw code is already better than no code}

Publishing raw code is already better than not publishing code at all,
as that raw code can be assumed to have run successfully at least once.
To reproduce an experiment from raw code can be painful, 
yet way better than rewriting the same code from scratch.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Publishing hosted code}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Code is sometimes published alongside a paper, 
in the form of a hosted website.
This way of publishing code gives additional options,
both for the authors of that code, as well as its users.

\paragraph{Example of hosted code}

There are multiple websites for hosting code, with GitHub,
and GitLab being the most well-known.
GitHub is the most commonly used website to host code in general,
with millions of users.
GitHub users can create dedicated websites (called 'repositories')
to upload code. GitHub hosts that uploaded by and provides
a user interface for every visiter of such a website.
An example from the author is the website \cite{bbbq_article_issue_157},
that hosts part of the code for the paper \cite{bilderbeek2022transmembrane}.

Hosted code has many additional features, such as
a version-controlled history, continuous integration scripts
and automatically generated metrics.

\paragraph{Hosted code has a commit history}

Hosted code commonly has a commit history. For example, GitHub
uses a version control system called \verb|git| (hence its name).
In practice, this means that when a change is made to the code,
a new version is created. In the case that the change was harmful,
one can go back to an earlier version and continue again from there.

\paragraph{A commit history helps assess honest reporting}

A commit history helps to assess the honesty 
in the reporting of the code. 
Ideally, the full version history of the code 
has been generated, by hosting the code directly at its inception.
However, in some repositories, only the final version of the code has been
posted, which means all historical information is lost.
From a code history, it can be observed, for example, 
how much statistical tests have really/actually been done 
and see if all are reported:
not reporting a statistical tests is considered a Dubious Research Practice.

\paragraph{Continuous integration helps assess code quality}

\paragraph{Metrics help assess the honest division of labour}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Publishing a virtual environment}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@
